---
output:
  github_document:
    html_preview: false
---

# dhms

<!-- [![Travis-CI Build Status](https://travis-ci.org/tidyverse/hms.svg?branch=master)](https://travis-ci.org/tidyverse/hms) [![AppVeyor Build Status](https://ci.appveyor.com/api/projects/status/github/tidyverse/hms?branch=master&svg=true)](https://ci.appveyor.com/project/tidyverse/hms) [![Coverage Status](https://img.shields.io/codecov/c/github/tidyverse/hms/master.svg)](https://codecov.io/github/tidyverse/hms?branch=master) [![CRAN_Status_Badge](http://www.r-pkg.org/badges/version/hms)](https://cran.r-project.org/package=hms) -->

A simple class for storing durations values and displaying them in the `Dd HH:MM:SS` format. 
The implementation is exhaustively based on the very neat [hms class](https://github.com/tidyverse/hms).
Intended to facilitate representation and manipulations of duration in the context of ethological experiments (that typically last several days).

The values are simply stored as a numeric vector a number of seconds. 
Implicitly, either since the onset of the experiment or since [ZT0](https://en.wikipedia.org/wiki/Zeitgeber) on the reference day.

```{r,eval=FALSE}
library(devtools)
install_github("qgeissmann/dhms")
library(dhms)
```

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
```






# Differences with `hms`

```{r}
library(dhms)
library(hms)
```

Display
-----------

In `dhms`, we express number of days to reduce cognitive burden after 72h.
This is convenient for circadian experiments:

```{r}
seconds <- seq(from =1, to =10 * 24 * 3600, by=27*3600)
df <- data.frame(hms=hms(seconds), dhms=dhms(seconds))
print(df)
```

Parsing
-----------

Parsing optional **number of days** and **negative values**:

```{r}
time_str <- c("12:34:56",         # regular format
              "12:34:56.789",     # decimal points
              "20:12",             # HH:MM (no seconds)
              "-12:34:56.001",    # negative values without days
              "1d 12:34:56",      # number of days (integer)
              "1d 12:34:56.001",  # decimal points and days
              "-1d 12:34:56.001", # negative values
              "-1d 12:34:56.001", # negative values and space
              "0.5d"             # day only
              
              )
df <- data.frame(time_str=time_str, hms=as.hms(time_str), dhms=as.dhms(time_str))
print(df)
```



Arithmetics
-----------

In `hms`, arithmetic operation return `difftime` object

```{r}
a <- hms(123) # 123 seconds
b <- hms(125)
print(a - b) # this returns a `difftime`
class(a - b)
print(a - 1) # this too
class(a - 1)
```

With `dhms`, we want to add or subtract constant to our time, but **keep expressing the result as a pretty duration**. Typically, you want to change the origin of your time, so you just subtract a number of seconds.
```{r}
a <- dhms(123)
b <- dhms(125)
print(a - b) # This returns another dhms, which expresses a duration
print(class(a - b))
print(a - 1) # Same here
print(class(a - 1))
```

Operations with time strings
---------------------------------------

In `hms`, we can't compare strings directly to time. 
One would have to use `as.hms()` on strings before comparison, which explicitly converts them to `hms` objects.
It is quite convenient to compare formatted strings directly to objects.
This is what happends in R for things like `POSIXct` objects:

```{r}
date = as.POSIXct("2017-12-12") 
print(date)
print(date  == "2017-12-12")
print(date  > "2015-06-12" )
```
Likewise, in `dhms`, operations with strings implicitly converts them string to a duration representation.

```{r}
time_str <- "00:02:10"
a_hms <- as.hms(time_str)
b_hms <- as.hms("00:02:02") 

a_dhms <- as.dhms(time_str)
b_dhms <- as.dhms("00:02:02") 

dt <- data.frame(operation = c("a == 00:02:10", "b > a", "b > 00:02:10", "a + 00:02:10"),
           hms=c(a_hms == time_str, b_hms > a_hms, b_hms > time_str, "FAILS"),
           dhms=c(a_dhms == time_str, b_dhms > a_dhms, b_dhms > time_str, as.character(a_dhms + time_str))
           )
print(dt)
```

