---
output:
  github_document:
    html_preview: false
---

# dhms
<!-- [![Travis-CI Build Status](https://travis-ci.org/tidyverse/hms.svg?branch=master)](https://travis-ci.org/tidyverse/hms) [![AppVeyor Build Status](https://ci.appveyor.com/api/projects/status/github/tidyverse/hms?branch=master&svg=true)](https://ci.appveyor.com/project/tidyverse/hms) [![Coverage Status](https://img.shields.io/codecov/c/github/tidyverse/hms/master.svg)](https://codecov.io/github/tidyverse/hms?branch=master) [![CRAN_Status_Badge](http://www.r-pkg.org/badges/version/hms)](https://cran.r-project.org/package=hms) -->

A simple class for storing durations values and displaying them in the `Dd HH:MM:SS` format. 
The implementation is exhaustivelly based on the very kneat [hms class](https://github.com/tidyverse/hms).
Intended to facilitate representation and manipulations of duration in the context of ethological expriments.

The values are stored as a numeric vector a number of seconds. 
Implicitly, either since the onset of the experiment, or [ZT0](https://en.wikipedia.org/wiki/Zeitgeber) on the reference day.
used in a data frame.



```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
```

```{r}
library(hms)
library(dhms)
```

|        function       |             hms        |          dhms       |
|:--------------------- | ----------------------:|:------------------- |
| `as.character()` |`hms(56, 34, 12, 2)` => `r as.character(hms(56, 34, 12, 2))` |  `dhms(56, 34, 12, 2)` => `r as.character(dhms(56, 34, 12, 2))` | 



